---
title: "R Demo"
author: "Jacob Hurley"
output:
  html_document:
    df_print: paged
---
```{r}
```
# Install library (enter into console)
```{r}
#install.packages('ggplot2')
#install.packages('dplyr')
#install.packges('tidyr')
```
# Load our library
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```
# Loading in our data
Kaggle Data is used for our initial demonstration -> https://www.kaggle.com/datasets/mrmorj/dataset-of-songs-in-spotify

Upload these files to the main project folder and call them here:
```{r}
#Loading our data into the R Global Environment
genre <- read.csv("genres_v2.csv", header = TRUE)
playlist <-read.csv("playlists.csv", header = TRUE)
```
# Head function
```{r}
head(genre)
```
# Tail Function
```{r}
tail(genre)
```
# Example of other data frame being selected
```{r}
head(playlist)
```
# 'str' used to return a list of objects and their structure
```{r}
str(genre)
```
# 'summary' gives us a brief summary of our data
```{r}
summary(genre)
```
# Brief graph demonstration using some r functions
```{r}
#Bar graphs
barplot(table(genre$energy))
```
```{r}
#a scatter plot between two variables
plot(genre$duration_ms, genre$valence)# type = "l" for a line graph
```

# Graphing with ggplot and dplyr
```{r}
genre %>% #dplyr command to load in data aka 'piping'
  mutate(duration_minutes = duration_ms/60000) %>% #more piping
  ggplot(aes(x = duration_minutes, y = energy)) + 
  ggtitle("Energy in a song vs Time (minutes)") +
  geom_smooth() +
  scale_x_continuous(breaks=seq(0,15,1))
```
